Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-02-06 paper.qmd
Updated Figure 17

tsne_plot_mnist_clust <- ggplot() + 
  geom_point(data = tsne_minst_scaled_temp,
             aes(
               x = tSNE1,
               y = tSNE2,
               colour = cluster
             ),
             alpha=0.2)+
  scale_color_manual(values=c('#f0f0f0', '#ff7f00')) +
  interior_annotation("a", sc_ltr_pos_mnist) +
  theme(legend.position = "none")

@@

Another usability of our approach is how NLDR methods behave with different density of the data structure. The tSNE layout of simulated $4\text{-}D$ C-shaped data with different densities are used. The high \pD{} model errors are scattered to a corner of the \gD{} layout of dense C-shaped structure. In the dense C-shaped structure, the high model errors occur because there are fewer data points positioned within the bins at the sparse end of the structure.

@@

The digit 1 of **Hand-written digits** data shows how the model accurately capture the non-linear structure. Not only that, the \pD{} model error help to find the anomalies of digit 1 images. Also, twisted pattern of the model evident that how tSNE tries to maintain the local structure by positioning the similar patterns of digit 1 close to each other.

@@

There are many avenues for future work. First, the development of
evaluation metrics is important for NLDR. We introduced a qualitative approach. But having a quantitative measure also will be useful for various application domains. Second, one could use our approach to create an approach for tuning (hyper-)parameters.  Another direction for future work is to introduce approaches to diagnose the fitted model. Finally, one can think of designing a new NLDR method by following the binning in high-dimensions.

>>>>

tsne_plot_mnist_clust <- ggplot() + 
  geom_point(data = tsne_minst_scaled_temp,
             aes(
               x = tSNE1,
               y = tSNE2,
               colour = cluster
             ),
             alpha=0.2)+
  scale_color_manual(values=c('#d9d9d9', '#ff7f00')) +
  interior_annotation("a", sc_ltr_pos_mnist) +
  theme(legend.position = "none")

@@

Another usability of our approach is how NLDR methods behave with different density of the data structure. The tSNE layout of simulated $4\text{-}D$ C-shaped data with different densities are used. The high $4\text{-}D$ model errors are scattered to a corner of the \gD{} layout of dense C-shaped structure. In the dense C-shaped structure, the high model errors occur because there are fewer data points positioned within the bins at the sparse end of the structure.

@@

The digit 1 of **Hand-written digits** data shows how the model accurately capture the non-linear structure. Not only that, the $10\text{-}D$ model error help to find the anomalies of digit 1 images. Also, twisted pattern of the model evident that how tSNE tries to maintain the local structure by positioning the similar patterns of digit 1 close to each other.

Diagnose the fitted model is important. Interactivity helps to

@@

There are many avenues for future work. First, the development of
evaluation metrics is important for NLDR. We introduced a qualitative approach. But having a quantitative measure also will be useful for various application domains. Second, one could use our approach to create an approach for tuning (hyper-)parameters.  Another direction for future work is to introduce more approaches to diagnose the fitted model. Finally, one can think of designing a new NLDR method by following the binning in high-dimensions.
